\documentclass[letterpaper, 11pt]{article}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{times}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage{graphicx}

\begin{document}

\title{Homework \#1 \\ Introduction to Algorithms/Algorithms 1 \\ 600.363/463 \\ Spring 2016}
\author{\textbf{Due on:} Thursday, February 4rd, 5pm \\
\textbf{Late submissions:} will NOT be accepted\\
\textbf{Format:} Please start each problem on a new page.
\\\textbf{Where to submit:} On blackboard, under student assessment. \\
Otherwise, please bring your solutions to the lecture.
\\}
\maketitle

\section{Problem 1 (25 points)}

\subsection{(15 points)}
For each statement below explain if it is true or false and prove
your answer. Be as precise as you can. The base of $\log$ is $2$ unless stated otherwise.

\begin{enumerate}

\item  ${n\over \log n} = \Theta(n)$

\item $2^n = o(3^{n})$

\item $2^n = \Theta(\left(\frac{3}{2}\right)^{n})$

\item $2^n = O(e^{n+\log n})$

\item $\log \log n = O(\log({\sqrt{n}\over 3\log n})+\log(n-\log n))$

\item Let $f,g,h$ be positive functions. Then $h(n)/(f(n)+g(n)) = O(h(n)/(f(n)g(n)))$

\item Let $f,g$ be positive functions. Then $|f(n) - g(n)| = \Theta(\min(f(n),g(n)))$

\item Let $f,g$ be positive functions. Then $f(n) + g(n) = \Theta(\max(f(n),g(n)))$

\item $n^{n\log n} = \Omega(e^{n^2 - n\log n})$

\item Let $f, g, h$ be a positive functions. If $g(n) = o(f(n))$ and $f(n) = O(h(n))$, then $g(n) = o(h(n))$.


\end{enumerate}


\subsection{(10 points)}

\begin{enumerate}

\item Prove that $$\sum_{i=1}^n i^2\ln\left(e + {1\over i}\right)  = \Theta(n^3).$$

\end{enumerate}

\section{Problem 2(25 Points)}

\subsection{(10 points)}

Prove by induction that ${2n \choose n} > \frac{2^n}{{\sqrt{\pi n}}}$ for all positive $n$.

\subsection{(15 points)}

\begin{enumerate}

\item Let $A,B,C$ be sets. Prove that 
\begin{enumerate}
\item$A\setminus (A \setminus B) = A \cap B$
\item$A \cup B = (A \setminus B) \cup (A \cap B) \cup (B \setminus A)$
\item$A \setminus (B \setminus C) = (A \setminus B) \cup (A \cap C)$
\end{enumerate}

\item There are $n$ identical apples in the basket. What is the number of different ways to divide the apples between Alice, Bob and John?
What is the answer if we request that Alice and John get at most $k < n/2$ apples each?


\item Suppose that a random machine outputs each number from $1$ to $x-1$ with equal probability. What is probability that the output is coprime with $x$, where $x = 3^n5^m7^k$ and $n,m,k \in \mathcal Z$. 

\item Assume there are $n$ students in a class. How many different ways are there to form $n/k$ teams with $k$ students in each?

\end{enumerate}
\end{document}